Density dependence

### 2020-11-12 analysis/paper/paper.Rmd
updated citation style and worked on outline

---
title: "Title Goes Here"
author:
  - Eric R. Scott:
      email: scottericr@gmail.com
      institute: [UF]
      correspondence: true
  - Paulo Rubim:
      email: rubimpaulo@gmail.com
      institute: [UFSC]
      correspondence: false
  - Maria Uriarte:
      email: mu2126@columbia.edu
      institute: [Columbia]
      correspondence: false
  - Emilio M. Bruna:
      email: embruna@ufl.edu
      intitute: [UF]
      correspondence: false

institute:
  - UF: Department of Wildlife Ecology and Conservation, University of Florida, Gainesville, Florida 32611-0430
  - UFSC: Departamento do Botânica, Universidade Federal de São Carlos, CP 676, CEP 13565-905, São Carlos, SP, Brazil
  - Columbia: Department of Ecology, Evolution and Environmental Biology, Columbia University 1200 Amsterdam Avenue, New York, New York 10027 USA
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  density dependence; vital rates; *Heliconia acuminata*
highlights: |
  These are the highlights
---

@@

@@

*Heliconia acuminata* (Heliconiaceae) is a perennial, self-incompatible understory monocot native to the nonflooded rainforests of central Amazonia. *Heliconia* are rhizomatious, but clonal reproduction via underground runners is rare in *H. acuminata*, unlike other congeners [@bruna2004]. Vegetative shoots grow from the rhizome as well as one or more flowering shoots when plants reproduce. Each flowering shoot has a single inflorescence of 20--25 flowers [@bruna2002]. Flowers are pollinated nearly exclusively by two species of hummingbirds at this site <!--# species? -->. Pollinated ovaries develop into fleshy fruits containing a maximum of 3 seeds which are dispersed exclusively by birds [@uriarte2011].

@@

## Data collection

The exact coordinates (to the nearest 0.1m) of each plant relative to the plot were recorded for a subset of plots in three years <!--# methods for collecting these data? -->. In 2006 and 2007, surveys were conducted to \_\_\_\_\_\_\_\_\_ . In 2009 a similar survey was conducted to assess the effects of safe-site limitation on recruitment [@uriarte2010]. For each plot and year, we calculated the density of conspecific shoots in annuli at varying distances from focal plants, as in Teller et al. [-@teller2016]. Density was calculated for 5 annuli ranging from an inner radius of 0m and an outer radius of 0.5m (i.e. a circle of radius 0.5m) to an inner radius of 4.5m and and outer radius of 5m. If an annulus intersected with the edge of the plot, the density for that particular plant and annulus was considered as a missing value. Conspecific density was estimated as the number of *H. acuminata* shoots per m^2^ in each annulus. Number of shoots was used as a measure of plant size rather than height as it is less error prone to measure, is not affected by plants being knocked over, and is strongly correlated with plant height and leaf area in *H. acuminata* [@bruna2002; @brunaEffectsForestFragmentation2002].

We used the three transition years for which we had density data to model the effects of conspecific density on survival, growth, and reproduction of *H. acuminata*. The collection of demographic data is described elsewhere [DATA PAPER], but briefly...

@@

<!-- Here's some example analysis code: -->

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r demo-plot, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```

Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot

```{r demo-inline-code}
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

@@

# Acknowledgements

<!-- The following line inserts a page break  -->

>>>>

---
title: "Interactive effects of drought and habitat fragmentation on vital rates of an understory tropical plant"
author:
  - Eric R. Scott:
      email: scottericr@gmail.com
      institute: [UF]
      correspondence: true
  - Paulo Rubim:
      email: rubimpaulo@gmail.com
      institute: [UFSC]
      correspondence: false
  - Maria Uriarte:
      email: mu2126@columbia.edu
      institute: [Columbia]
      correspondence: false
  - Emilio M. Bruna:
      email: embruna@ufl.edu
      intitute: [UF]
      correspondence: false

institute:
  - UF: Department of Wildlife Ecology and Conservation, University of Florida, Gainesville, Florida 32611-0430
  - UFSC: Departamento do Botânica, Universidade Federal de São Carlos, CP 676, CEP 13565-905, São Carlos, SP, Brazil
  - Columbia: Department of Ecology, Evolution and Environmental Biology, Columbia University 1200 Amsterdam Avenue, New York, New York 10027 USA
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/ecology.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  forest fragmentation; vital rates; *Heliconia acuminata*; drought; ENSO; edge effects
highlights: |
  These are the highlights
---

@@

## Habitat fragmentation

Habitat fragmentation is bad, pervasive, and getting worse

-   stats about global land use change <!--# start this broad or start with the Amazon? -->

-   stats about amazon deforestation [@malhi2008]

-   stats about how much land is near an edge (refs in Schwartz et al and betts 2019)

-   Why fragmentation bad

    -   microclimate changes [@ewersFragmentationImpairsMicroclimate2013]

    -   metapopulation and connectivity stuff (necessary?)

-   Plants in the tropics are less well adapted to fragmentation because lack of historical distrubance and less variable climate [@betts2019]

## Climate change and drought

-   The northern Amazon has been experiencing a drying trend since the mid 1970s and ensemble climate models predict substantial decreases in dry season precipitation in southern Amazonia for the 21st century [@malhi2008].

-   The El Niño--Southern Oscillation (ENSO) reduces wet-season rainfall in the Amazon <!--# is this correct? Or only in the south? -->. ENSO-induced droughts are predicted to increase in frequency and severity. [@caiIncreasingFrequencyExtreme2014].

-   Also sea surface temperature related drought. Warming of north Atlantic results longer and more intense dry season in southern and eastern Amazonia [@phillipsDroughtSensitivityAmazon2009; @malhi2008]

-   drought effects on plant vital rates

-   1997 ENSO drought [@williamsonAmazonianTreeMortality2000]

-   2005 ENSO drought [@phillipsDroughtSensitivityAmazon2009; @marengoDroughtAmazonia20052008]

-   2010 drought [@lewis2010AmazonDrought2011]

-   [@zengCausesImpacts20052008]

## Interactions

-   However, our understanding of how climate change and habitat fragmentation might interact to impact organisms is limited.

-   [@schwartz2019] (mention how tree size classes affected differently, but we don't know if this is also true for understory herbaceous plants.

-   [@malhi2008]

-   The 1997--1998 ENSO drought caused greater tree mortality and increased leaf litter fall in experimentally fragmented forests compared to intact forest in the Amazon [@laurancePositiveFeedbacksForest2001].

-   Furthermore, the interactive effects of fragmentation and drought may be difficult to detect in long-lived species (e.g. Heliconia) [bruna paper that shows growth is reduced after 10 years slightly, but impacts reproduction]

## This experiment

-   The "fragmentation drought hypothesis" expects drought to have stronger impacts on forest fragments since fragments less buffered from changes in humidity and more sensitive to desiccation. Under this hypothesis we would expect drought to reduce growth, survival, reproduction and recruitment more in fragmented landscapes compared to intact forest.

-   It's also possible that plants in fragments have already been filtered such that the survivors are already adapted to more variable and desiccating conditions associated with forest edges [@schwartz2019]. This may result in drought having a less noticeable impact on plants in fragments, although we would expect fragmentation to still have an impact on survival, growth, and reproduction.

-   We set out to answer the following questions: 1) Does drought have different impacts on plants in forest fragments compared to intact forest? 2) Do the magnitudes and directions of these effects differ among vital rates? 3) Do the effects of drought, fragmentation, and their interaction vary with plant size [@tredennickSizebyenvironmentInteractionsNeglected2018] <!--# read this referrebce --> ?

@@

*Heliconia acuminata* (Heliconiaceae) is a perennial, self-incompatible understory monocot native to the nonflooded rainforests of central Amazonia. *Heliconia* are rhizomatious, but clonal reproduction via underground runners is rare in *H. acuminata*, unlike other congeners [@bruna2004]. Vegetative shoots grow from the rhizome as well as one or more flowering shoots when plants reproduce. Each flowering shoot has a single inflorescence of 20--25 flowers [@bruna2002]. Flowers are pollinated nearly exclusively by two species of hummingbirds at this site <!--# species? -->. Pollinated ovaries develop into fleshy fruits containing a maximum of 3 seeds which are dispersed exclusively by birds [@uriarte2011; @brunaHeliconiaAcuminataSeed2014].

@@

@@

@@

# Acknowledgments

### 2020-11-12 analysis/supplementary-materials/appendix_density_dependence.Rmd
updated citation style and worked on outline

---
title: "Density dependence"
author: "Eric R. Scott"
date: '2020-09-13'
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

@@

>>>>

---
title: "Density dependence"
author: "Eric R. Scott"
date: '2020-09-13'
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/ecology.csl" # Insert path for the bib-style
---

@@

-   Cite [@bruna2004] (reproductive success is independent of local floral density. Not sure if scales are comparable though, need to read paper first)
